---
title: "Beer Breweries Data"
author: "Sanjee Yogeswaran"
date: "June 26, 2017"
output:
  html_document:
    keep_md: yes
---

## Introduction

This is a case study on breweries and their ABV and IBU content. We explore the data for beer breweries around the United States, and particularly focus on alcoholic content and international bitterness units for these breweries. 

The number of breweries in each state are as follows:
```{r}
## Code to read in GDP and Education tables
beers <- read.csv("data/beers.csv", header=TRUE)
breweries <- read.csv("data/breweries.csv",header=TRUE)
## Question 1:
##Code to count number of breweries in each state
table(breweries$State)
```
The two tables are merged together, and the first six and last six entries are shown.
```{r}
## Question 2:
## Code to merge the two tables together
## Display first 6 rows and last 6 rows
mergedtables <- merge(beers,breweries, by.x=("Brewery_id"),by.y = ("Brew_ID"))
head(mergedtables,6)
tail(mergedtables,6)
```
The number of NAs in each column are displayed below:
```{r}
## Question 3:
##Count number of NAs in each column
colSums(is.na(mergedtables))
```
The median alcohol content and IBU of each state are shown below, along with bar graphs of each for comparison:
```{r}
## Question 4:
##Compute median alcohol content
abv <- tapply(mergedtables$ABV,mergedtables$State,median,na.rm=T)
abv
##Compute median IBU
ibu <- tapply(mergedtables$IBU,mergedtables$State,median,na.rm=T)
ibu
##Plot bar graphs for ABV and IBU
library(ggplot2)
barplot(abv)
barplot(ibu)
```

The maximum values for ABV and IBU for each state and their respective values are printed below:
```{r}
## Question 5:
##Max values for ABV and IBU
which.max(abv)
max(abv)
which.max(ibu)
max(ibu,na.rm=T)
```
The summary statistics for ABV is printed here:
```{r}
## Question 6:
##Summary statistics for ABV
summary(abv)
```
The relationship between bitterness of beer and alcoholic content is explored here. A scatter plot is created as well as a linear regression model.
```{r}
##  Question 7:
##Plot abv vs ibu
qplot(abv,ibu)

```

There is a weak general positive trend shown in the scatter plot. The linear regression below explores this further.
```{r}
##Regression model
summary(lm(ibu ~ abv, data=mergedtables))
cor.test(ibu, abv)
```
The scatter plot shows a weak linear relationship between alcoholic beverage content and international bitterneess units. The adjusted r-squared value of 0.0604 reinforces that it is a weak correlation. The p-value generated in testing the linear model is 0.04, which shows that there is evidence to reject the null hypothesis that there is no correlation, however, since the value is close to 0.05, it shows it is a weak relationship.

##Conclusion
The data from brewieries in each state are analyzed in this case study. Two datasets are mergeda together and the total number of breweries in each state are displayed. The first six and last six entries in the merged tables are then displayed. Then the number of non available entries are shown. The median alcohol content and international bitterness units are calculated and plotted on a bar graph. The maximal values for the alcohol content and international bitterness units are found by state and their respective values. Then the summary statistics for alcoholic content is displayed. Finally a relationship between the international bitterness units and alcoholic content is explored through a scatter plot and a linear relationship.
